\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2019-01-26 Documents/M1/Astronum_2018/sections/Conclusion.tex

We have developed IMEX schemes suitable for a two-moment model of neutrino transport in \texttt{thornado} that can obey Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin method for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, some schemes having convex combinations as their stages should be able to preserve the convexity.
This idea encourages us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and eventually a method with realizability-preserving IMEX time integrator and high-order DG method.
See \cite{chu_etal_2018} for some examples with third-order DG methods.

In the application motivating this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
(We also run the code with other algebraic closures for examination.
And the program crashed after few steps because of the nonphysical neutrino density/flux.)

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have strong-stability preserving property in the streaming limit.  
Their accuracy were demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and elastic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combined an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration can be a candidate for constructing a CCSN simulation code without "nonphysical protection" that may break conservation laws.

In this work, we adopted Cartesian coordinates, linear collision term, and a fixed non-relativistic material background.
More realistic problems of scientific interest, such as with the scattering with energy exchange and relativistic effects, are left for future research.

>>>>

We have developed IMEX schemes suitable for a two-moment model of neutrino transport in \texttt{thornado} that obeys Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin method for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, some schemes having convex combinations as their stages should be able to preserve the realizable domain.
This idea encourages us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and eventually a method with a realizability-preserving IMEX time integrator and high-order DG method.
See~\cite{chu_etal_2018} for more details and examples of the third-order DG methods.

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
(We also ran the code with other algebraic closures for examination.
And the program crashed after a few time steps because of the nonphysical neutrino density/flux obtained and no correction step was taken for $f>1$.
We could make the code run by adding correction step, but the quality of the results may be compromised.
And it's difficult to determine whether the results are in fact compromised by a particular correction step and, if so, in what way and to what degree.)

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have the strong-stability preserving property in the streaming limit.  
Their accuracy was demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and isoenergetic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combining an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration can be a candidate for constructing a CCSN simulation code without "unphysical correction steps" that may break conservation laws.

In this work, we adopted Cartesian coordinates, a linear collision term, and a fixed non-relativistic material background.
More realistic problems of scientific interest, such as with energy-exchanging scattering and relativistic effects, are left for future research.

### 2019-01-26 Documents/M1/Astronum_2018/sections/SpatialDiscretization.tex

In this section we discretize the two-moment model with a simple first-order finite volume method to illustrate how the closure affects the realizability-preserving property of the method.  
By assuming that the moments at time level $t^{n}$ ($\bcM^{n}$) satisfy the bounds in Eq.~\eqref{eq:MomentsBounds}, our goal is to identify sufficient conditions such that the moments at time level $t^{n+1}$ ($\bcM^{n+1}$) also satisfy the bounds.  
To simply the illustration, we limit the discussion to one spatial dimension and employ a uniform Cartesian mesh.  
(The extension to multiple spatial dimensions and high-order discretization using the discontinuous Galerkin method is given in \cite{chu_etal_2018})

@@

Consider Eq.~\eqref{eq:MomentIMEX} and assume that $\bcM_{i}^{n}$ is realizable for all $i$.  
Lemma~3 in \cite{chu_etal_2018} states that $\bcM^{n+1}_{i}$ is realizable provided $\f{\dt}{\tau} > 0$ and $\widetilde{\bcM}^{n}_{i}$ is realizable.  
In Eq.~\eqref{eq:widetildeM}, if $\beta \in [0,1]$, $\widetilde{\bcM}^{n}_{i}$ is expressed as a convex combination of $\bcM_{i}^{n}$ and the expression in the square brackets on the right-hand side of Eq.~\eqref{eq:widetildeM}.  
It follows that $\widetilde{\bcM}^{n}_{i}$ is realizable if the expression inside the square brackets is realizable.  
By Lemma~2 in \cite{chu_etal_2018}, the expression in square brackets is realizable for a distribution satisfying $f\in[0,1]$.  
For the two-moment model considered here, realizability depends in the algebraic closure.  
Specifically, if $\bcM_{i}^{n}$ is realizable and the Eddington factor satisfies the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}, then $\bcM^{n+1}_{i}$ is realizable provided $\beta \in [0,1]$.  
Thus, realizability of $\bcM^{n+1}_{i}$ requires both a closure based on Fermi-Dirac statistics and a CFL condition $\dt\le\dx$.

>>>>

In this section we discretize the two-moment model with a simple first-order finite volume method to illustrate how the closure affects the realizability-preserving property of the method.  
By assuming that the moments at time level $t^{n}$ ($\bcM^{n}$) satisfy the bounds in Eq.~\eqref{eq:MomentsBounds}, our goal is to identify sufficient conditions such that the moments at time level $t^{n+1}$ ($\bcM^{n+1}$) also satisfy the bounds.  
To simply the illustration, we limit the discussion to one spatial dimension and employ a uniform Cartesian mesh.  
(The extension to multiple spatial dimensions and high-order discretization using the discontinuous Galerkin method is given in~\cite{chu_etal_2018}.)

@@

Consider Eq.~\eqref{eq:MomentIMEX} and assume that $\bcM_{i}^{n}$ is realizable for all $i$.  
Lemma~3 in~\cite{chu_etal_2018} states that $\bcM^{n+1}_{i}$ is realizable provided $\f{\dt}{\tau} > 0$. 
%and $\widetilde{\bcM}^{n}_{i}$ is realizable.  
In Eq.~\eqref{eq:widetildeM}, if $\beta \in [0,1]$, $\widetilde{\bcM}^{n}_{i}$ is expressed as a convex combination of $\bcM_{i}^{n}$ and the expression in the square brackets on the right-hand side of Eq.~\eqref{eq:widetildeM}.  
It follows that $\widetilde{\bcM}^{n}_{i}$ is realizable if the expression inside the square brackets is realizable.  
By Lemma~2 in~\cite{chu_etal_2018}, the expression in square brackets is realizable for a distribution satisfying $f\in[0,1]$.  
For the two-moment model considered here, realizability depends on the algebraic closure.  
Specifically, if $\bcM_{i}^{n}$ is realizable and the Eddington factor satisfies the bounds in Eq.~\eqref{eq:eddingtonFactorBounds}, then $\bcM^{n+1}_{i}$ is realizable provided $\beta \in [0,1]$.  
Thus, realizability of $\bcM^{n+1}_{i}$ requires both a closure based on Fermi-Dirac statistics and a CFL condition $\dt\le\dx$.

### 2019-01-26 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Work on astronum paper

\clearpage
\section{Time Integration} \label{se:TimeIntegration}

Suppose that an algebraic closure based on Fermi-Dirac statistics is used (i.e., the Eddington factor satisfies Eq.~\eqref{eq:eddingtonFactorBounds}).
Here we consider the construction of an Implicit-Explicit (IMEX) time integration scheme which maintains the bounds in Eq.~\eqref{eq:MomentsBounds}.  
The semi-discretization of the two-moment model results in a system of ordinary differential equations of the form
\begin{equation}
  \dot{\vect{u}} = \vect{\cT}(\vect{u}) + \vect{\cQ}(\vect{u}),
\end{equation}
where the solution vector
\begin{equation}
  \vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-averaged moments, $\vect{\cT}$ is the transport operator, corresponding to the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\vect{\cQ}$ is the collision operator, corresponding to the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

Since the set of realizable moments is convex, convex-invariant schemes, which maintain states in a convex set, can be used to design realizability-preserving methods for the two-moment model. 
Ideally, the scheme should also be high-order accurate and work well in the asymptotic diffusion limit (characterized by frequent collisions and long time scales).  
The following discussion considers the construction of such convex-invariant schemes.

@@

Treating the transport operator explicitly and the collision operator implicitly, a standard $s$-stage IMEX scheme takes the following form \cite{pareschiRusso_2005}: 
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where $(\tilde{a}_{ij})$ and $(a_{ij})$, coefficients of the $i$-th stage, are elements of matrices $\tilde{A}$ and $A$, respectively.
The matrices $\tilde{A}$ and $A$ are lower triangular ($\tilde{A}$ is strictly lower triangular so that the transport part is explicit).  
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weights in the assembly step in Eq.~\eqref{imexIntermediate}.
These coefficients and weights must satisfy certain order conditions for consistency, accuracy, and other properties.  
For second-order temporal accuracy, the following conditions are required \cite{hairer_1981}:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j=1}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j=1}^{s}a_{ij}$.

The IMEX scheme is called globally stiffly accurate (GSA) if the coefficients satisfy \cite{dimarcoPareschi2013}:
\begin{equation}
  a_{si}=w_{i} \quad\text{and}\quad \tilde{a}_{si}=\tilde{w}_{i}, \quad \text{for} \quad i=1,\ldots,s.
\end{equation}
Then, $\vect{u}^{n+1} = \vect{u}^{(s)}$, which is simplifying because the assembly step in Eq.~\eqref{imexIntermediate} is omitted.  
IMEX schemes are further classified by the structure of implicit matrix $A$.  
If $A$ is invertible, the IMEX scheme is of type~A \cite{pareschiRusso_2005}.  
If $a_{i1} = 0$ for $i=1,\ldots,s$, $w_{1} = 0$, and the submatrix consisting of the last $s-1$ rows and columns is invertible, the IMEX scheme is of type~ARS \cite{ascher_etal_1997,pareschiRusso_2005}.

@@

If the IMEX scheme is GSA, $\vect{u}^{n+1} = \vect{u}^{(s)}$.  
Moreover, if $c_{ij},\tilde{c}_{ij}\ge0$ and $a_{ii}>0$, each stage in Eq.~\eqref{eq:imexStagesRewrite} is a convex combination of explicit Euler steps (with time step $\hat{c}_{ij}\dt$), followed by a implicit Euler step.  
Each of the explicit Euler steps has a time step condition that ensures its realizability given by $\hat{c}_{ij}\,\dt\leq\dx$; the Courant-Friedrichs-Lewy (CFL) condition of the scheme.
Using results proved in \cite{chu_etal_2018} and discussed in Section~\ref{se:SpacialDiscretization}, the IMEX scheme is convex-invariant and realizability-preserving for the two-moment model in Section~\ref{se:SpacialDiscretization} provided
\begin{equation}
  \max(\hat{c}_{ij})\,\dt \leq \dx.  
\end{equation}
(This CFL condition becomes more restrictive with high-order DG spatial discretization \cite{chu_etal_2018}.)

@@

Accuracy in the diffusion limit is another important property to consider for the IMEX scheme applied to the two-moment model.  
In the diffusion limit, the distribution function is nearly isotropic so that $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$ and $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, and the two-moment model is approximately governed by (e.g., \cite{jinLevermore_1996})
\begin{equation}
  \pd{\cJ}{t} + \nabla\cdot\vect{\cH} = 0
  \quad\text{and}\quad
  \vect{\cH} = - \tau\,\nabla\cdot\vect{\cK}.  
  \label{eq:diffusionLimit}
\end{equation}
In the context of IMEX schemes, the above relationships imply that the following relations should hold
\begin{equation}
   \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=1,\ldots,s,
   \label{diffusionAccuracy}
\end{equation}
where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix and $\vect{e}$ is the vector of ones.
Eq.~\eqref{diffusionAccuracy} implies
\begin{equation}
  c_{i} = \tilde{c}_{i}, \quad i=1,\ldots,s
\end{equation}
We have proved in \cite{chu_etal_2018} that only IMEX schemes of type~ARS can be both diffusion accurate and convex-invariant.  
(Another short proof follows from the fact that IMEX schemes type~A have $\tilde{c}_1 = 0$ while $c_i \neq 0$.)

@@

Unfortunately, coefficients satisfying the order conditions in Eqs.~\eqref{orderConditions1}-\eqref{orderConditions2} and the conditions for convex-invariance do not exist for the standard IMEX scheme in Eqs.~\eqref{imexStages}-\eqref{imexIntermediate}, unless a small time step is invoked that makes the scheme essentially explicit.  
To circumvent this problem, correction steps can be introduced after the assembly step in Eq.~\eqref{imexIntermediate} (e.g., \cite{chertock_etal_2015,hu_etal_2018}).  
However, the correction steps can impose time step constraints for realizability or accuracy in the diffusion limit that ruins the efficiency gains expected from the IMEX scheme.  
Because of this, we sacrifice overall high-order accuracy, and aim for IMEX schemes that are high-order accurate in the streaming limit, diffusion accurate, and convex-invariant.  
Combining these requirements we seek GSA IMEX schemes of type~ARS with coefficients satisfying the following constraints:
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item High-order accuracy in the streaming limit.  For second-order accuracy: \ee{(Do we need to list third-order accuracy also?)}
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2}.
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Diffusion accurate:
    \begin{equation}
      c_{i}=\tilde{c}_{i}, \quad i=1,\ldots,s.
      \label{eq:diffusionCondition}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
      \label{eq:convexInvariant}
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}:=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    \item Less than five stages ($s\le4$) \label{cod:statges}.
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. 
\end{enumerate}
Fortunately, these IMEX schemes are easy to find.  
(The constraint in \eqref{cod:statges} is introduced from efficiency considerations to limit the number of implicit solves.)
We call the IMEX schemes satisfying the above conditions {PD-ARS} (see also Definition~3 in \cite{chu_etal_2018}), and we provide two optimal PD-ARS schemes below: PD-ARS with SSPRK2 and PD-ARS with SSPRK3, which limit to the optimal second-order and third-order SSPRK schemes from \cite{shuOsher_1988}, respectively.

@@

The optimal PD-ARS with SSPRK2 in the standard double Butcher tableau form, explicit tableau on the left ($\tilde{A}$) and implicit tableau on the right ($A$), is given by:
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array}
\end{equation}
For this scheme, only two implicit solves are needed per time step and $c_{\mbox{\tiny Sch}}= 1$, which implies that the time step restriction for preserving moment realizability is only due to the explicit part.

@@

The optimial PD-ARS with SSPRK3 is given in its standard double Butcher tableau form (explicit tableau on the left and implicit tableau on the right) by:
\begin{equation}
  \begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}
\end{equation}
This scheme requires three implicit solvers per time step, and $c_{\mbox{\tiny Sch}}= 1$.  
Since PD-ARS with SSPRK3 is not more accurate than PD-ARS with SSPRK2 in collision dominated regions (see results in Section~\ref{se:NumericalTests}), it may not offer any practical advantage over PD-ARS with SSPRK2.

>>>>

\section{Time Integration} \label{se:TimeIntegration}

Suppose that an algebraic closure based on Fermi-Dirac statistics is used (i.e., the Eddington factor satisfies Eq.~\eqref{eq:eddingtonFactorBounds}).
Here we consider the construction of an Implicit-Explicit (IMEX) time integration scheme that maintains the bounds in Eq.~\eqref{eq:MomentsBounds}.  
The semi-discretization of the two-moment model results in a system of ordinary differential equations of the form
\begin{equation}
  \dot{\vect{u}} = \vect{\cT}(\vect{u}) + \vect{\cQ}(\vect{u}),
\end{equation}
where the solution vector
\begin{equation}
  \vect{u}(t) = \left( \bcM_{1}(t),\ldots,\bcM_{N}(t)\right) ^{T}
\end{equation}
is the collection of all cell-averaged moments, $\vect{\cT}$ is the transport operator, corresponding to the first term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, and $\vect{\cQ}$ is the collision operator, corresponding to the second term on the right-hand side of Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}.

Since the set of realizable moments is convex, convex-invariant schemes, which preserve the initial value's convex domain, can be used to design realizability-preserving methods for the two-moment model. 
Ideally, the scheme should also be high-order accurate and work well in the asymptotic diffusion limit (characterized by frequent collisions and long time scales).  
The following discussion considers the construction of such convex-invariant schemes.

@@

Treating the transport operator explicitly and the collision operator implicitly, a standard $s$-stage IMEX scheme takes the following form~\cite{pareschiRusso_2005}: 
\begin{align}
  \vect{u}^{(i)}
  &=\vect{u}^{n}
  +\dt\sum_{j=1}^{i-1}\tilde{a}_{ij}\,\vect{\cT}(\vect{u}^{(j)})
  +\dt\sum_{j=1}^{i}a_{ij}\,\vect{\cQ}(\vect{u}^{(j)}),
  \quad i=1,\ldots,s, \label{imexStages} \\
  \vect{u}^{n+1}
  &=\vect{u}^{n}
  +\dt\sum_{i=1}^{s}\tilde{w}_{i}\,\vect{\cT}(\vect{u}^{(i)})
  +\dt\sum_{i=1}^{s}w_{i}\,\vect{\cQ}(\vect{u}^{(i)}), \label{imexIntermediate} 
\end{align}
where $(\tilde{a}_{ij})$ and $(a_{ij})$, coefficients of the $i$-th stage, are elements of matrices $\tilde{A}$ and $A$, respectively.
The matrices $\tilde{A}$ and $A$ are lower triangular ($\tilde{A}$ is strictly lower triangular so that the transport part is explicit).  
The vectors $\tilde{\vect{w}}=(\tilde{w}_{1},\ldots,\tilde{w}_{s})^{T}$ and $\vect{w}=(w_{1},\ldots,w_{s})^{T}$ are the weights in the assembly step in Eq.~\eqref{imexIntermediate}.
These coefficients and weights must satisfy certain order conditions for consistency, accuracy, and other properties.  
For second-order temporal accuracy, the following conditions are required~\cite{hairer_1981}:
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=\sum_{i=1}^{s}w_{i}=1,
  \label{orderConditions1}
\end{equation}
and
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}\tilde{w}_{i}\,c_{i}
  =\sum_{i=1}^{s}w_{i}\,\tilde{c}_{i}
  =\sum_{i=1}^{s}w_{i}\,c_{i}=\f{1}{2}, 
  \label{orderConditions2}
\end{equation}
where $\tilde{c}_{i} = \sum_{j=1}^{s}\tilde{a}_{ij}$ and $c_{i}=\sum_{j=1}^{s}a_{ij}$.

The IMEX scheme is called globally stiffly accurate (GSA) if the coefficients satisfy~\cite{dimarcoPareschi2013}:
\begin{equation}
  a_{si}=w_{i} \quad\text{and}\quad \tilde{a}_{si}=\tilde{w}_{i}, \quad \text{for} \quad i=1,\ldots,s.
\end{equation}
Then, $\vect{u}^{n+1} = \vect{u}^{(s)}$, which is simplifying because the assembly step in Eq.~\eqref{imexIntermediate} is omitted.  
IMEX schemes are further classified by the structure of the implicit matrix $A$.  
If $A$ is invertible, the IMEX scheme is of type~A~\cite{pareschiRusso_2005}.  
If $a_{i1} = 0$ for $i=1,\ldots,s$, $w_{1} = 0$, and the submatrix consisting of the last $s-1$ rows and columns is invertible, the IMEX scheme is of type~ARS~\cite{ascher_etal_1997,pareschiRusso_2005}.

@@

If the IMEX scheme is GSA, $\vect{u}^{n+1} = \vect{u}^{(s)}$.  
Moreover, if $c_{ij},\tilde{c}_{ij}\ge0$ and $a_{ii}>0$, each stage in Eq.~\eqref{eq:imexStagesRewrite} is a convex combination of explicit Euler steps (with time step $\hat{c}_{ij}\dt$), followed by a implicit Euler step.  
Each of the explicit Euler steps has a time step condition that ensures its realizability given by $\hat{c}_{ij}\,\dt\leq\dx$; the Courant-Friedrichs-Lewy (CFL) condition of the scheme.
Using results proved in~\cite{chu_etal_2018} and discussed in Section~\ref{se:SpacialDiscretization}, the IMEX scheme is convex-invariant and realizability-preserving for the two-moment model in Section~\ref{se:SpacialDiscretization} provided
\begin{equation}
  \max(\hat{c}_{ij})\,\dt \leq \dx.  
\end{equation}
(This CFL condition becomes more restrictive with high-order DG spatial discretization~\cite{chu_etal_2018}.)

@@

Accuracy in the diffusion limit is another important property to consider for the IMEX scheme applied to the two-moment model.  
In the diffusion limit, the distribution function is nearly isotropic so that $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$ and $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, and the two-moment model is approximately governed by (e.g., \cite{jinLevermore_1996})
\begin{equation}
  \pd{\cJ}{t} + \nabla\cdot\vect{\cH} = 0
  \quad\text{and}\quad
  \vect{\cH} = - \tau\,\nabla\cdot\vect{\cK}.  
  \label{eq:diffusionLimit}
\end{equation}
In the context of IMEX schemes, the above relationships imply that the following relations should hold~\cite{chu_etal_2018}
\begin{equation}
   \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=1,\ldots,s,
   \label{diffusionAccuracy}
\end{equation}
where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix, $\vect{e}$ is the vector of ones, and $A$ and $\tilde{A}$ are the matrices of the coefficients $(\tilde{a}_{ij})$ and $(a_{ij})$.
Eq.~\eqref{diffusionAccuracy} implies
\begin{equation}
  c_{i} = \tilde{c}_{i}, \quad i=1,\ldots,s
\end{equation}
We have proved in \cite{chu_etal_2018} that only IMEX schemes of type~ARS can be both diffusion accurate and convex-invariant.  
(Another short proof follows from the fact that IMEX schemes of type~A have $\tilde{c}_1 = 0$ while $c_i \neq 0$.)

@@

Unfortunately, coefficients satisfying the order conditions in Eqs.~\eqref{orderConditions1}-\eqref{orderConditions2} and the conditions for convex-invariance do not exist for the standard IMEX scheme in Eqs.~\eqref{imexStages}-\eqref{imexIntermediate}, unless a small time step is invoked that makes the scheme essentially explicit.  
To circumvent this problem, correction steps can be introduced after the assembly step in Eq.~\eqref{imexIntermediate} (e.g., \cite{chertock_etal_2015,hu_etal_2018}).  
However, the correction steps can impose time step constraints for realizability or accuracy in the diffusion limit that ruins the efficiency gains expected from the IMEX scheme.  
Because of this, we sacrifice overall high-order accuracy, and aim for IMEX schemes that are high-order accurate in the streaming limit, diffusion accurate, and convex-invariant.  
Combining these requirements we seek GSA IMEX schemes of type~ARS with coefficients satisfying the following constraints~\cite{hu_etal_2018,chu_etal_2018}:
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item High-order accuracy in the streaming limit.
    For second-order accuracy:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2}.
      \label{eq:orderConditionsEx}
    \end{equation}
    For third-order accuracy: 
    \begin{equation}
    \sum_{i=1}^{s}\tilde{w}_{i}=1,
          \quad
          \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
          \quad
          \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c_{i}}^2 = \f{1}{3}
          \quad\text{and}\quad
          \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{a_{ij}}\tilde{c}_{j}= \f{1}{6}
    \end{equation}
    \item Diffusion accurate:
    \begin{equation}
      c_{i}=\tilde{c}_{i}, \quad i=1,\ldots,s.
      \label{eq:diffusionCondition}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
      \label{eq:convexInvariant}
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}:=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    \item Less than five stages ($s\le4$) \label{cod:statges}.
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. 
\end{enumerate}
Fortunately, these IMEX schemes are easy to find.  
(The constraint in \eqref{cod:statges} is introduced from efficiency considerations to limit the number of implicit solves.)
We call the IMEX schemes satisfying the above conditions {PD-ARS} (see also Definition~3 in~\cite{chu_etal_2018}), and we provide two optimal PD-ARS schemes below: (PD-ARS with SSPRK2) and (PD-ARS with SSPRK3), which limit to the optimal second-order and third-order SSPRK schemes from~\cite{shuOsher_1988}, respectively.

@@

The optimal PD-ARS with SSPRK2 in the standard double Butcher tableau form, explicit tableau on the left ($\tilde{A}$) and implicit tableau on the right ($A$), is given by:
\begin{align}
  &\begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0 
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array} \\ 
  &\hspace{48pt} \tilde{A} \hspace{90pt} A
\end{align}
For this scheme, only two implicit solves are needed per time step and $c_{\mbox{\tiny Sch}}= 1$, which implies that the time step restriction for preserving moment realizability is only due to the explicit part.

@@

The optimal PD-ARS with SSPRK3 is given in its standard double Butcher tableau form (explicit tableau on the left and implicit tableau on the right) by:
\begin{align}
  &\begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}\\ 
    &\hspace{60pt} \tilde{A} \hspace{130pt} A
\end{align}
This scheme requires three implicit solves per time step, and $c_{\mbox{\tiny Sch}}= 1$.  
Since (PD-ARS with SSPRK3) is not more accurate than (PD-ARS with SSPRK2) in collision dominated regions (see results in Section~\ref{se:NumericalTests}), it may not offer any practical advantage over (PD-ARS with SSPRK2).
